An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-05 EconGrowth.Rnw
Now doing U predictions.

**************** To Do List ********************
\begin{itemize}
\item Reese: Add text from Word verson of paper to the LaTeX version.
\item \st{Heun: Create a table of AIC values for each fit.}
\item Heun: Add $u$ preductions for Cobb-Douglas
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the Cobb-Douglas with energy rows 
      \begin{itemize}
      \item ZA: lower bound on $\lambda$ and upper bound on $\alpha$
      \item ZM: lower bound on $\lambda$
      \end{itemize}
\item Heun: Fix warnings of the form ``Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562'' in the code that generates the Cobb-Douglas with energy fits.
\item Heun: Add CES production function fits
\item Heun: Add Linex production function fits
\item Heun: Add $k$, $l$, $q$, $x$, and $u$ fits ONLY
\end{itemize}
*************************************************

@@

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(tikzDevice) #See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  warning=TRUE, #Tells whether to show warnings.
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (cache=TRUE).
  cache=TRUE,
  echo=FALSE,    #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

@@

#########################
# Takes the Cobb-Douglas fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cobbDouglasPredictions <- function(countryAbbrev, energyType){
  model <- cobbDouglasModel(countryAbbrev, energyType)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # There are two countries which do not have labor data for 2011. Need to add a "NA" for 2011 for those countries.
  if (countryAbbrev == "CN" || countryAbbrev == "ZA"){
    dfNA <- data.frame(NA)
    colnames(dfNA) <- "pred" #Need to give it the same name so that we can apply rbind
    df <- rbind(df, dfNA)
  }
  return(df)
}

@@

##############################
# Creates a graph that plots predicted GDP as lines and GDP data as open circles.
##
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the CD model without energy
  predictionsQ <- cobbDouglasPredictionsColumn("Q") #Predictions with Q
  predictionsX <- cobbDouglasPredictionsColumn("X") #Predictions with X
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictions, predictionsQ, predictionsX) 
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "red", "blue") #line and symbol colors
  symbols <- c(1, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP+predGDPQ+predGDPX ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "No energy", "With $q$", "With $x$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ), 
                  ylim=list(c(1,8), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3)),
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
@

@@

\[ a \in[0,1], b \in [0,1], \alpha=\min(a,b), \beta=|b-a|, \gamma = 1-\max(a,b) \]

@@

<<AIC_Analysis>>=
###############################
# Creates a xtable object that holds the AIC values for each parameter estimation that we include.
##
createAICTable <- function(){
  # Cobb-Douglas
  models <- lapply(countryAbbrevs, cobbDouglasModel)
  aicCD <- data.frame(lapply(models, AIC))
  rownames(aicCD) <- "CD"
  # Cobb-Douglas with Q
  models <- lapply(countryAbbrevs, cobbDouglasModel, energyType="Q")
  aicCDq <- data.frame(lapply(models, AIC))
  rownames(aicCDq) <- "CD$q$"
  # Cobb-Douglas with X
  models <- lapply(countryAbbrevs, cobbDouglasModel, energyType="X")
  aicCDx <- data.frame(lapply(models, AIC))
  rownames(aicCDx) <- "CD$x$"
  # Create the data.frame table
  out <- rbind(aicCD, aicCDq, aicCDx) 
  # Create the xtable with caption and label
  out <- xtable(out, 
                caption="AIC values for all models.", 
                label="tab:AICTable", 
                digits=1 #Only one number after the decimal place
                )
  return(out)
}
@

>>>>

**************** To Do List ********************
\begin{itemize}
\item Reese: Add text from Word verson of paper to the LaTeX version.
\item \st{Heun: Create a table of AIC values for each fit.}
\item \st{Heun: Add $u$ predictions for Cobb-Douglas}
\item Heun: Add covariance metrics.
\item Heun: Eliminate blanks in the coefficient tables for the 95\% CIs in the Cobb-Douglas with energy rows. Asked Pruim about this via email but have not heard a response.
      \begin{itemize}
      \item ZA: lower bound on $\lambda$ and upper bound on $\alpha$
      \item ZM: lower bound on $\lambda$
      \end{itemize}
\item Heun: Fix warnings of the form ``Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562'' in the code that generates the Cobb-Douglas with energy fits.
\item Heun: Add CES production function fits
\item Heun: Add Linex production function fits
\item Heun: Add $k$, $l$, $q$, $x$, and $u$ ONLY fits
\end{itemize}
*************************************************

@@

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
require(tikzDevice) #See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  warning=TRUE, #Tells whether to show warnings.
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (cache=TRUE).
  cache=FALSE,
  echo=FALSE,    #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

@@

#########################
# Takes the Cobb-Douglas fitted models and creates per-country predictions for them.
# Returns a data.frame with the predictions.
##
cobbDouglasPredictions <- function(countryAbbrev, energyType){
  # Can't make predictions for any of CN, ZA, SA, IR, TZ, or ZM if we're interested in U
  if (!missing(energyType)){
    # Consider this replacement only if energyType has been specified.
    if (energyType == "U" && (countryAbbrev == "CN" || countryAbbrev == "ZA" || countryAbbrev == "SA" 
                              || countryAbbrev == "IR" || countryAbbrev == "TZ" || countryAbbrev == "ZM")){
      #Return a column of NAs if the above conditions have been met.
      nRows <- 21 # All of these countries need 21 rows.
      df <- as.data.frame(matrix(NA, ncol = 1, nrow = nRows))
      colnames(df) <- "pred"
      return(df)
    }
  }
  model <- cobbDouglasModel(countryAbbrev, energyType)
  pred <- predict(model) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
  df <- data.frame(pred)
  # Pad with rows as necessary
  df <- padRows(countryAbbrev, df)
  return(df)
}

#####################
# This function adds NA rows to a data.frame to ensure that the df is the same length as the country data set
##
padRows <- function(countryAbbrev, df){
  nRowsData <- nrow(loadData(countryAbbrev)) #Gets number of rows in data set
  nRowsDF <- nrow(df)
  nRowsToAdd <- nRowsData - nRowsDF
  nCol <- ncol(df)
  dfToAppend <- as.data.frame(matrix(NA, ncol=nCol, nrow=nRowsToAdd))
  colnames(dfToAppend) <- colnames(df)
  return(rbind(df, dfToAppend))
}

@@

##############################
# Creates a graph that plots predicted GDP as lines and GDP data as open circles.
##
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn()     #Predictions from CD without energy
  predictionsQ <- cobbDouglasPredictionsColumn("Q") #Predictions from CD with Q
  predictionsX <- cobbDouglasPredictionsColumn("X") #Predictions from CD with X
  predictionsU <- cobbDouglasPredictionsColumn("U") #Predictions from CD with U
  #Now add the predictions columns to the data.
  dataTable <- cbind(dataTable, predictions, predictionsQ, predictionsX, predictionsU) 
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP+predGDPQ+predGDPX+predGDPU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("Actual", "No energy", "With $q$", "With $x$", "With $u$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ), 
                  ylim=list(c(1,8), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3), c(1,3)),
                  xlab="", ylab="Indexed GDP (1980=1 or 1991=1)") #axis labels
  return(graph)
}
@

@@

$a \in[0,1], b \in [0,1], \alpha=\min(a,b), \beta=|b-a|, \gamma = 1-\max(a,b)$

@@

<<AIC_Analysis>>=
###############################
# Creates an xtable object that holds the AIC values for each parameter estimation that we include.
##
createAICTable <- function(){
  # Cobb-Douglas
  models <- lapply(countryAbbrevs, cobbDouglasModel)
  aicCD <- data.frame(lapply(models, AIC))
  rownames(aicCD) <- "CD"
  # Cobb-Douglas with Q
  models <- lapply(countryAbbrevs, cobbDouglasModel, energyType="Q")
  aicCDq <- data.frame(lapply(models, AIC))
  rownames(aicCDq) <- "CD$q$"
  # Cobb-Douglas with X
  models <- lapply(countryAbbrevs, cobbDouglasModel, energyType="X")
  aicCDx <- data.frame(lapply(models, AIC))
  rownames(aicCDx) <- "CD$x$"
  # Cobb-Douglas with U
  models <- c(cobbDouglasModel("US", "U"), cobbDouglasModel("UK", "U"), cobbDouglasModel("JP", "U"))
  aicCDu <- cbind(US=AIC(cobbDouglasModel("US", "U")), 
                  UK=AIC(cobbDouglasModel("UK", "U")), 
                  JP=AIC(cobbDouglasModel("JP", "U")),
                  CN=NA, ZA=NA, SA=NA, IR=NA, TZ=NA, ZM=NA) #No U data for these countries.
  rownames(aicCDu) <- "CD$u$"
  # Create the data.frame table
  out <- rbind(aicCD, aicCDq, aicCDx, aicCDu) 
  # Create the xtable with caption and label
  out <- xtable(out, caption="AIC values for all models.", label="tab:AICTable", digits=1)
  return(out)
}
@
